---
title: "Working with NFL Official Stats"
output: 
  html_notebook:
    theme: united
    highlight: tango
    tidy: styler
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##### Load packages
```{r load, message=FALSE, warning=FALSE}
library("nflfastR")
library("skimr")
library("gt")
library("tidyverse")
```

```{r import, message=FALSE, include=FALSE}
# raw data is pulled from data folder
rosters <- read_csv("./data/rosters.csv") %>% arrange(-season, team, position, gsis_id)
games <- read_csv("./data/games.csv")
snaps <- read_csv("./data/snaps.csv")
raw_weekly_stats <- read_csv("./data/weekly_stats.csv")
```

```{r join include=FALSE}
joined_weekly_stats <- raw_weekly_stats %>%
  left_join(rosters %>% select(-team), by = c("player_id" = "gsis_id", "season")) %>%
  left_join(games, by = c("season", "week", "recent_team" = "team")) %>%
  left_join(snaps %>% select(-recent_team), by = c("player_id", "season", "week"))
```

Season stats
```{r aggregate include=FALSE}
# define variables to be summarized
sum_cols <- joined %>%
  select_if(is.numeric) %>%
  select(
    -contains(c("_id", "_rate", "_pct")),
    -c(season, week, years_exp:spread_line, adot)
  ) %>%
  colnames()

mean_cols <- joined %>%
  select_if(is.numeric) %>%
  select(-c(season, week, years_exp:spread_line, started_bi:over_half_bi)) %>%
  colnames()

# season
raw_season_stats <- joined_weekly_stats %>%
  group_by(season, player_id) %>%
  summarise(
    full_name = last(full_name),
    recent_team = last(recent_team),
    position = last(position),
    nw_position = last(nw_position),
    games = n(),
    across(all_of(sum_cols),
      sum,
      na.rm = TRUE
    )
  )

# per game
raw_per_game_stats <- joined_weekly_stats %>%
  group_by(season, player_id) %>%
  summarise(
    across(all_of(mean_cols),
      mean,
      .names = "{.col}_pg",
      na.rm = TRUE
    )
  )

# join per game stats to aggregate stats
joined_season_stats <- raw_season_stats %>%
  left_join(raw_per_game_stats, by = c("player_id", "season"))

# tidy and add fantasy ranks
season_stats <- joined_season_stats %>%
  group_by(season, position) %>%
  arrange(-nw_pts) %>%
  mutate(nw_rk = 1:n()) %>%
  group_by(season) %>%
  mutate(nw_rk_ovr = 1:n()) %>%
  ungroup() %>%
  rename(
    games_started = started_bi,
    games_active = active_bi,
    games_healthy = healthy_bi,
    games_over_half = over_half_bi
  ) %>%
  relocate(c(nw_position, player_id), .after = last_col()) %>%
  mutate(across(where(is.numeric), round, 2)) %>%
  arrange(-season, nw_rk_ovr) %>%
  select(-c(contains("fumbles"), sack_yards))
```

```{r}
glimpse(season_stats)
```

```{r variables}
season_select <- season_stats %>%
  select(season, full_name, position, recent_team, games) %>% colnames()
  colnames()
weekly_select <- joined %>%
  select(season, week, full_name, position, recent_team) %>%
  colnames()
nw_cols <- joined %>%
  select(contains(c("yards_", "tds_", "_pts")), -passing_yards_after_catch, -receiving_yards_after_catch) %>%
  colnames()
passing_cols <- season_stats %>%
  select(completions:int_tds) %>%
  colnames()
rushing_cols <- season_stats %>%
  select(carries:rushing_tds_50, touches) %>%
  colnames()
receiving_cols <- season_stats %>%
  select(receptions:receiving_tds_50) %>%
  colnames()
universal_cols <- season_stats %>% select(opportunities, 
                                          
                                          nw_pts, fantasy_points, nw_rk, std_rk, nw_rk_ovr, nw_position, player_id) %>% colnames()
passing_pg_cols <- season_stats %>%
  select(completions_pg:int_tds_pg) %>%
  colnames()
rushing_pg_cols <- season_stats %>%
  select(carries_pg:rushing_tds_50_pg) %>%
  colnames()
receiving_pg_cols <- season_stats %>%
  select(receptions_pg:receiving_tds_50_pg) %>%
  colnames()
```


## Datasets {.tabset .tabset}

### Rosters

The 'rosters' dataset provides fantasy relevant player data and ids of from 2010 to 2020. This dataset is primarily used to gather positional data to join with stats. The ids provided may also be utilized to further join outside data.
```{r rosters}
rosters
```


### Games

The 'games' dataset provides game data and results from 2010 to 2020.
```{r games}
games
```

### Weekly

The 'weekly' dataset provides weekly player stats, snaps, injury data, and further advanced metrics that will be added down the road. Each observation is unique by a player_id and game_id that unites the season, week, home team, and away team.

```{r}
weekly
```

### Season

The 'season' dataset aggregates the 'weekly' dataset by season.

```{r}
season
```

## {-}

## Usage 


### Quarterbacks

```{r}
qbs <- season_stats %>%
  filter(position == "QB") %>%
  select(nw_rk, season, full_name, position, recent_team, passing_cols, player_id) %>%
  arrange(-season, nw_rk)
```


### Running Backs
```{r}
rbs <- season_stats %>%
  filter(position == "RB") %>%
  select(nw_rk, season, full_name, position, recent_team, rushing_cols, player_id) %>%
  arrange(-season, nw_rk)
```


### Receivers
```{r}
receivers <- season_stats %>%
  filter(position %in% c("WR", "TE")) %>%
  select(nw_rk, season, full_name, position, recent_team, receiving_cols, player_id) %>%
  arrange(-season, nw_rk)
```


##### 2021 Season

```{r}
raw_fp_ranks <- fp_rankings(page = "consensus-cheatsheets")
fp_ranks <- raw_fp_ranks %>%
  select(rank:worst, contains("_id")) %>%
  mutate(fantasypros_id = as.numeric(fantasypros_id))
```

```{r}
fantasy_21 <- rosters_offense %>%
  filter(season == 2021, !is.na(team)) %>%
  select(season, full_name, nw_position, depth_chart_position, team, years_exp, status, fantasypros_id, gsis_id) %>%
  left_join(fp_ranks) %>%
  arrange(rank) %>%
  rename(player_id = gsis_id)

qbs <- fantasy_21 %>%
  filter(nw_position == "QB") %>%
  left_join(season_stats[id_cols, passing_cols])
```


```{r}
# 2021 coaches
games[1:32, ] %>% select(season, team, team_coach)
```


```{r}
season %>% left_join(
  nw_rosters_20 %>%
    rename(mfl_id = player_id) %>%
    mutate(season = 2020) %>%
    select(mfl_id, season, franchise_name, drafted) %>%
    mutate(mfl_id = as.numeric(mfl_id)),
  by = c("season", "mfl_id")
)
```
